### 2017-03-06 README.Rmd
updated README and switched to non-standard evaluation in autoplot

pdp: An R Package for Constructing Partial Dependence Plots
================

The primary purpose of this package is to provide a general framework for constructing _partial dependence plots_ (PDPs) in R.

### Installation

The R package `pdp` is available from [CRAN](https://cran.r-project.org/package=pdp); the development version is hosted on [GitHub](https://github.com/bgreenwell/pdp). There are two ways to install:
```{r, eval=FALSE}
# Install latest release from CRAN (recommended)
install.packages("pdp")

# Install development version from GitHub repo
devtools::install_github("bgreenwell/pdp")
```

### Example usage

The examples below demonstrate various usages of the `pdp` package: regression, classification, and interfacing with the well-known `caret` package. To start, we need to install a few additional packages that will be required to run the examples.
```{r, eval=FALSE}
install.packages(c("caret", "ggmap", "kernlab", "randomForest"))
```

#### Regression example

In this example, we fit a random forest to the Boston housing data. (See `?boston` for a brief explanation of the data.) Note, for any of the following examples, you can see a progress bar by simply specifying `progress = "text"` in the call to `partial`. You may also reduce the computation time via the `grid.resolution` option in the call to `partial`.
```{r, message = FALSE, warning=FALSE, fig.width=12, fig.height=4}
# Load required packages
library(pdp)  # for constructing PDPs
library(randomForest)  # for random forest algorithm

# Load the Boston housing data
data (boston)  # included with the pdp package

# Fit a random forest to the boston housing data
set.seed(101)  # for reproducibility
boston.rf <- randomForest(cmedv ~ ., data = boston)

# Partial dependence of lstat and rm on cmedv
grid.arrange(
  partial(boston.rf, pred.var = "lstat", plot = TRUE, rug = TRUE),
  partial(boston.rf, pred.var = "rm", plot = TRUE, rug = TRUE),
  partial(boston.rf, pred.var = c("lstat", "rm"), plot = TRUE, chull = TRUE),
  ncol = 3
)

```

#### Classification example

In this example, we fit a support vector machine with a radial basis function kernel to the Pima Indians diabetes data. (See `?pima` for a brief explanation of the data.)
```{r, message=FALSE, warning=FALSE, fig.width=12, fig.height=4}
# Load required packages
library(kernlab)  # for fitting SVMs

# Fit an SVM to the Pima Indians diabetes data
data (pima)  # load the boston housing data
pima.svm <- ksvm(diabetes ~ ., data = pima, type = "C-svc", kernel = "rbfdot",
                 C = 0.5, prob.model = TRUE)

# Partial dependence of glucose and age on diabetes test result (neg/pos).
grid.arrange(
  partial(pima.svm, pred.var = "glucose", plot = TRUE, train = pima),
  partial(pima.svm, pred.var = "age", plot = TRUE, train = pima),
  partial(pima.svm, pred.var = c("glucose", "age"), plot = TRUE, train = pima),
  ncol = 3
)
```

#### Interface with `caret`

Finally, we demonstrate the construction of PDPs from models fit using the `caret` package; `caret` is an extremetly useful package for classification and regression training that, essentially, has one function (`train`) for fitting all kinds of predictive models in R (e.g., `glmnet`, `svm`, `xgboost`, etc.).

For illustration we use `caret`'s `train` function to tune an [XGBoost](https://github.com/dmlc/xgboost) model to the Pima Indians diabetes data using 5-fold cross-validation. We then use the final model to construct PDPs for `glucose` and `age`. Note, when training a model using `caret`'s `train` function, you can view tuning progress by setting `verboseIter = TRUE` in the call to `trainControl`.
```{r, message=FALSE, warning=FALSE, fig.width=15, fig.height=5}
# Load required packages
library(caret)  # for model training/tuning

# Set up for 5-fold cross-validation
ctrl <- trainControl(method = "cv", number = 5)

# Grid of tuning parameter values
xgb.grid <- expand.grid(
  nrounds = c(1000, 2000),
  max_depth = 2:4,
  eta = c(0.001, 0.01, 0.1),
  gamma = 0, 
  colsample_bytree = 1,
  min_child_weight = 1,
  subsample = c(0.5, 0.75, 1)
)

# Tune am XGBoost model to the Pima Indians diabetes data. This may take a few 
# minutes!
set.seed(103)  # for reproducibility
pima.xgb <- train(diabetes ~ ., data = pima, method = "xgbTree",
                  prob.model = TRUE, na.action = na.omit, trControl = ctrl,
                  tuneGrid = xgb.grid)

# Partial dependence of glucose and age on diabetes test result (neg/pos)
grid.arrange(
  partial(pima.xgb, pred.var = "glucose", plot = TRUE, rug = TRUE),
  partial(pima.xgb, pred.var = "age", plot = TRUE, rug = TRUE),
  partial(pima.xgb, pred.var = "mass", plot = TRUE, rug = TRUE),
  ncol = 3 
)
```

>>>>

# pdp: An R Package for Constructing Partial Dependence Plots

Complex nonparametric models---like neural networks, random forests, and support vector machines---are more common than ever in predictive analytics, especially when dealing with large observational databases that don't adhere to the strict assumptions imposed by traditional statistical techniques (e.g., multiple linear regression which assumes linearity, homoscedasticity, and normality). Unfortunately, it can be challenging to understand the results of such models and explain them to management. Partial dependence plots offer a simple solution. Partial dependence plots are low-dimensional graphical renderings of the prediction function $\widehat{f}\left(\boldsymbol{x}\right)$ so that the relationship between the outcome and predictors of interest can be more easily understood. These plots are especially useful in explaining the output from black box models. The `pdp` package offers a general framework for constructing partial dependence plots for various types of fitted models in R.

A detailed introduction to `pdp` has been accepted for publication in The R Journal; a preprint is available [here](https://github.com/bgreenwell/pdp-paper/blob/master/RJwrapper.pdf). You can track development on  at https://github.com/bgreenwell/pdp. To report bugs or issues, contact the main author directly or submit them to https://github.com/bgreenwell/pdp/issues.

As of right now, `pdp` exports four functions:

* `partial` - compute partial dependence functions (i.e., objects of class `"partial"`) from various fitted model objects;
* `plotPartial"` - plot partial dependence functions (i.e., objects of class `"partial"`) using `lattice` graphics;
* `autoplot` - plot partial dependence functions (i.e., objects of class `"partial"`) using `ggplot2` graphics;
* `topPredictors` extract most "important" predictors from various types of fitted models.

## Installation

The `pdp` package is [currently listed on CRAN](https://CRAN.R-project.org/package=pdp) and can easily be installed:
```r
  # Install from CRAN (recommended)
  install.packages("pdp")

  # Alternatively, install the development version from GitHub
  devtools::install_github("bgreenwell/pdp")
```
